---
title: "Report"
author: "Ido Ali"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Investing in Indonesia's energy sector is really important because since it helps the country grow economically, creates jobs, and powers up industries. But Indonesia has some challenges in keeping up with its energy needs, so investing locally is key to making sure Indonesia have enough energy and don't rely too much on imports.

When it comes to picking which energy company to invest in, you've gotta look at a few things. Check out how well the company is doing financially, see if the people running it know what they're doing, and make sure they're not wasting resources. You also need to consider to go for companies that care about the environment and are trying to use cleaner energy sources.

Before you invest in a company, it's important to check if they're doing well financially. You want to make sure they're making money and are steady. There are different things you can look at to figure this out.

Today, we'll look at how these companies are doing financially by checking their earnings, P/E, P/B, and P/S ratios. The companies we're looking at are:

1. Adaro Energy Indonesia Tbk
2. Bayan Resources Tbk
3. Dian Swastatika Sentosa Tbk
4. Golden Energy Mines Tbk
5. Indo Tambangraya Megah Tbk

# Import Data

To assess the companies, we need data. We'll use data we got from the Sectors App API. This API is handy for analyzing the finances of companies listed on the Indonesia Stock Exchange (IDX).

The API will become public soon.

```{r}
library(jsonlite)
library(dplyr)

stock_symbols <- c("ADRO.JK", "BYAN.JK", "DSSA.JK", "GEMS.JK", "ITMG.JK")
stock_names <- c("adaro", "bayan", "dian", "golden", "indo")
stock_fullnames <- c("Adaro Energy Indonesia Tbk",
                     "Bayan Resources Tbk",
                     "Dian Swastatika Sentosa Tbk",
                     "Golden Energy Mines Tbk",
                     "Indo Tambangraya Megah Tbk")

load_json_data <- function(symbol) {
  file_path <- paste0("data_input/", symbol, ".json")
  json_data <- fromJSON(file_path)
  return(json_data)
}

stock_data <- lapply(stock_symbols, load_json_data)
```

# Exploration

```{r}
library(showtext)

font <- "Gudea"
font_add_google(family=font, font, db_cache = TRUE)
fa_path <- systemfonts::font_info(family = "Font Awesome 6 Brands")[["path"]]
font_add(family = "fa-brands", regular = fa_path)

showtext_auto(enable = TRUE)
```

## Earnings

Examining earnings is crucial before making investment decisions because earnings reflect a company's profitability, which is a key indicator of its financial health and performance. Higher earnings generally indicate that a company is generating more profits, which can lead to increased shareholder value through dividends or reinvestment for future growth. Moreover, earnings growth over time is often seen as a positive sign, indicating the company's ability to expand its business and generate sustainable returns for investors.

```{r}
library(tidyr)

all_earnings <- data.frame(year = seq(2019, 2022))

for (i in seq(1, length(stock_symbols))){
  new <- stock_data[[i]]$historical_financials[, c("year", "earnings")] %>%
    mutate(!!stock_names[i] := earnings) %>% 
    select(year, stock_names[i])
  
  all_earnings <- all_earnings %>% 
    left_join(new, by = "year")
}

```

```{r}
library(ggplot2)
library(scales)
library(ggtext)

company_switch <- function(x){
  index <- match(x, stock_names)
  y <- stock_fullnames[index]
  return(y)
}

all_earnings %>% 
  pivot_longer(cols = stock_names, names_to = "company") %>% 
  mutate(company = company_switch(company)) %>% 
  ggplot(aes(x = year, y = value)) +
  geom_smooth(aes(color = company), size = 2) +
  geom_point(aes(color = company), size = 4) +
  scale_y_continuous(labels = label_number(suffix = " T", scale = 1e-12)) +
  scale_color_manual(values = colors) +
  labs(title = "<b>Earning Comparisons</b>",
       subtitle = "<b>Adaro</b> and <b>Bayan</b> have the highest earnings<br>out of all<br><i>(All earnings are in Rupiahs)</i>") +
  # theme_void() +
  theme(legend.position = "bottom",
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(size = 1, color = txt_col),
        plot.background = element_rect(fill = bg_col),
        panel.background = element_rect(fill = bg_col),
        legend.title = element_blank(),
        legend.background =  element_blank(),
        legend.key = element_rect(fill = bg_col),
        text = element_text(color = txt_col, family = "Gudea"),
        axis.text = element_text(color = txt_col, size = 12),
        axis.title = element_blank(),
        plot.title = element_markdown(size = 40, family = "Gudea", hjust = 0.5,
                                      margin = margin(0, 0, 10, 0)),
        plot.subtitle = element_markdown(size = 15, family = "Gudea",
                                         hjust = 0.5, margin = margin(0, 0, 20, 0))) +
  guides(color = guide_legend(nrow = 2))
  
```

Bayan and Adaro appear to be the most attractive investment options based on their consistently high earnings compared to the other companies in the oil & gas sector in Indonesia. 

Consistent and high earnings suggest that these companies have strong operational performance and potentially better prospects for future growth, making them more appealing to investors seeking stable returns.

## P/B 

The P/B ratio compares a company's market value to its book value (assets minus liabilities). It provides insights into whether a stock is undervalued or overvalued relative to its accounting value. A P/B ratio less than 1 may indicate that the stock is undervalued, while a ratio greater than 1 may suggest overvaluation. However, P/B ratio should be interpreted cautiously.

Let's see how P/B ratio comparisons between each companies.

```{r}
valuation <- data.frame()

for (i in seq(1, length(stock_names))){
  new_valuation <- stock_data[[i]]$historical_valuation %>% 
    filter(year == 2023) %>% 
    select(pb, pe, ps) %>% 
    mutate(company = stock_fullnames[i])
  
  valuation <- rbind(valuation, new_valuation)
}
```


```{r}
library(ggtext)

txt_col <- "#f2f2f2"
bg_col <- "#0c0a09"
colors <- c("#e11d48", "#62bb9f", "#574086", "#366e9c", "#d68b53")

valuation %>% 
  ggplot(aes(x = pb, y = reorder(company, pb))) +
  geom_col(aes(fill = company)) +
  scale_x_continuous(limits = c(0, 22)) +
  geom_textbox(x = 15, y = stock_fullnames[4],
            label = "<b>Bayan</b> is 4 times bigger\nthan <b>Dian</b>",
            fill = txt_col, color = bg_col) +
  geom_text(aes(x = pb + 1, y = reorder(company, pb), label = round(pb, 1)),
            color = txt_col) +
  scale_fill_manual(values = colors) +
  labs(title = "<b>Bayan</b> Is Overvalued",
       subtitle = "From P/B, We can See that Bayan has<br>P/B value bigger than <b>other 4 companies combined</b>") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.box.background = element_rect(fill = bg_col),
        panel.background = element_rect(fill = bg_col),
        plot.background = element_rect(fill = bg_col),
        text = element_text(color = txt_col, family = "Gudea"),
        plot.title = element_markdown(hjust = 0.5, size = 40),
        plot.subtitle = element_markdown(hjust = 0.5, size = 15,
                                         margin = margin(10, 0, 10, 0)),
        aspect.ratio = 0.3
        ) +
  guides(fill = guide_legend(nrow = 2))
  
```

As you can see above, Bayan Resources Tbk has P/B ratio higher than all other companies combined. 

## P/E

The P/E ratio indicates how much investors are willing to pay for each dollar of a company's earnings. A lower P/E ratio may suggest that a company is undervalued relative to its earnings, while a higher P/E ratio may indicate that it is overvalued.

However, a low P/E ratio alone does not necessarily mean a good investment, as it could be due to poor growth prospects or other underlying issues. 

That's why beside looking at P/E ratio, you also nee to look at other values. Like Earning for example. 

```{r}
valuation %>% 
  ggplot(aes(x = pe, y = reorder(company, pe))) +
  geom_col(aes(fill = company)) +
  scale_x_continuous(limits = c(0, 32)) +
  geom_textbox(x = 20, y = stock_fullnames[4],
            label = "<b>Bayan Resources Tbk</b> P/B ratio is far bigger than others",
            fill = txt_col, text.colour = bg_col, halign = 0.5) +
  geom_text(aes(x = pe + 1, y = reorder(company, pe), label = round(pe, 1)),
            color = txt_col) +
  scale_fill_manual(values = colors) +
  labs(title = "<b>P/B Ratio</b> Comparison",
       subtitle = "Same as P/B ratio, P/E ratio shows that<br><b>Bayan Resources Tbk is overvalued</b>") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.box.background = element_rect(fill = bg_col),
        panel.background = element_rect(fill = bg_col),
        plot.background = element_rect(fill = bg_col),
        text = element_text(color = txt_col, family = "Gudea"),
        plot.title = element_markdown(hjust = 0.5, size = 40),
        plot.subtitle = element_markdown(hjust = 0.5, size = 15,
                                         margin = margin(10, 0, 10, 0)),
        aspect.ratio = 0.3
        ) +
  guides(fill = guide_legend(nrow = 2))
```

Just like what we observed in P/B ratio, P/E ratio also shows that Bayan Resources Tbk is overvalued. 

## P/S Ratio

The P/S ratio compares a company's market capitalization to its total sales revenue. It helps investors assess the valuation of a company relative to its revenue generation. A lower P/S ratio may indicate undervaluation, while a higher ratio may suggest overvaluation.

However, similar to other metrics, it should be used in conjunction with other factors for a comprehensive analysis.

```{r}
valuation %>% 
  ggplot() +
  geom_col(aes(x = ps, y = reorder(company, ps), fill = company)) +
  scale_x_continuous(limits = c(0, 12)) +
  geom_textbox(x = 6, y = stock_fullnames[4],
            label = "<b>Bayan Resources Tbk</b> P/S ratio is far bigger than others",
            fill = txt_col, color = bg_col) +
  geom_text(aes(x = ps + 0.5, y = reorder(company, ps), label = round(ps, 1)),
            color = txt_col) +
  scale_fill_manual(values = colors) +
  labs(title = "<b>P/S Ratio</b> Comparison",
       subtitle = "Same as P/B ratio and P/E ratio<br>P/S ratio also shows that<br><b>Bayan Resources Tbk is overvalued</b>") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.box.background = element_rect(fill = bg_col),
        panel.background = element_rect(fill = bg_col),
        plot.background = element_rect(fill = bg_col),
        text = element_text(color = txt_col, family = "Gudea"),
        plot.title = element_markdown(hjust = 0.5, size = 40),
        plot.subtitle = element_markdown(hjust = 0.5, size = 15,
                                         margin = margin(10, 0, 10, 0)),
        aspect.ratio = 0.3
        ) +
  guides(fill = guide_legend(nrow = 2))
```

P/S ratio also shows the same thing as P/B ratio and P/E ratio. P/S ratio of Bayan Resources Tbk is still higher than all other companies P/S ratio combined times two. 

# Conclusion

It appears that Bayan Resources may be overvalued, especially when compared to Adaro Energy Indonesia. Despite both companies consistently boasting the highest and most stable earnings each year, their differing P/E ratio, P/B ratio, and P/S ratio suggest that the market views their values differently.

Therefore, if considering an investment among these five companies:
1. Adaro Energy Indonesia Tbk
2. Bayan Resources Tbk
3. Dian Swastatika Sentosa Tbk
4. Golden Energy Mines Tbk
5. Indo Tambangraya Megah Tbk

This report recommends investing in Adaro Energy Indonesia.

However, it's important to note that besides earnings and the aforementioned ratios, there are other critical factors to consider. Factors such as company culture, market trends, public policy, and others should also be taken into account when making investment decisions. 






